---
title: "ed_slow_convergence"
author: "Yunqi Yang"
date: "10/11/2023"
output: html_document
---

## Description:
Compare the convergence of TED vs. ED. We first run ED algorithm for 20 iterations, and then run TED/ED for 1e6 iterations. I did the experiment on 4 different data sets. The first two are data simulated with large n/R(n = 1e5, R = 5), the last two are data simulated from small n/R(n=1e3, R=50). 

TED stopped at a few hundreds or thousands iterations because the log-likelihood starts to decrease to a very small amount, -1.455192e-11 for all 4 data. Wondering if this is the smallest number that R can handle? I think it means TED converges there. 

ED will run 1e6 iterations. 


```{r}
library(udr)
library(mvtnorm)
```

```{r}
plot_likelihood <- function(loglik1, loglik2, names, cols, xlab, ylab, type, pch, cex){
  ymin <- min(loglik1, loglik2)
  ymax <- max(loglik1, loglik2)
  xmax <- max(length(loglik1), length(loglik2))
  plot(loglik1, col = cols[1], xlab = xlab, ylab = ylab, xlim = c(0, xmax + 2), ylim = c(ymin-2, ymax + 2), type, pch = pch, cex = cex)
  lines(loglik2, col = cols[2], pch = pch, cex = cex)
  legend("bottomright", legend = names, col = cols, pch = pch)
}

plot_convergence <- function(loglik1, loglik2, names, cols, xlab, ylab, pch, e, cex, lwd, type){
  fbest = max(max(loglik1), max(loglik2))
  dat1 <- data.frame(iter = 1:length(loglik1),
                   dist = fbest - loglik1 + e)  
  dat2 <- data.frame(iter = 1:length(loglik2),
                   dist = fbest - loglik2 + e)  
  ymin <- min(log10(dat1$dist), log10(dat2$dist))
  ymax <- -log10(e)
  xmax <- max(length(loglik1), length(loglik2))
  plot(dat1$iter, log10(dat1$dist), col = cols[1], ylim = c(ymin-0.5, ymax+0.5), 
       xlim = c(0, xmax + 2),  ylab = ylab, xlab = xlab, type = type, pch = pch, cex = cex, lwd = lwd)
  lines(dat2$iter, log10(dat2$dist), col = cols[2], pch = pch, cex = cex, type = type, lwd = lwd)
  legend("bottomright", legend = names, col = cols, pch = pch)
}
```




## Data example 1:
```{r}
ted1 <- readRDS("/Users/nicholeyang/Desktop/udr_related/udr_simulation202212/fig1_convergence/ted.rds")
ed1 <- readRDS("/Users/nicholeyang/Desktop/udr_related/udr_simulation202212/fig1_convergence/ed.rds")
```
## Data example 2:
```{r}
ted2 <- readRDS("/Users/nicholeyang/Desktop/udr_related/udr_simulation202212/fig1_convergence/ted3.rds")
ed2 <- readRDS("/Users/nicholeyang/Desktop/udr_related/udr_simulation202212/fig1_convergence/ed3.rds")
```

## Data example 3:
```{r}
ted3 <- readRDS("/Users/nicholeyang/Desktop/udr_related/udr_simulation202212/fig1_convergence/ted4.rds")
ed3 <- readRDS("/Users/nicholeyang/Desktop/udr_related/udr_simulation202212/fig1_convergence/ed4.rds")
```

## Data example 4:
```{r}
ted4 <- readRDS("/Users/nicholeyang/Desktop/udr_related/udr_simulation202212/fig1_convergence/ted8.rds")
ed4 <- readRDS("/Users/nicholeyang/Desktop/udr_related/udr_simulation202212/fig1_convergence/ed8.rds")
```





```{r}
pdf("/Users/nicholeyang/Desktop/udr_related/udr_simulation202306/Fig1/ed_slow_convergence.pdf", width = 7, height = 12)
par(mfrow = c(4,2), mar = c(7, 5, 2, 2))

# example 1 
ted <- data.frame(iter = 1: (nrow(ted1$progress)-20),
		  loglik = ted1$progress$loglik[21:nrow(ted1$progress)])
ed  <- data.frame(iter = 1:1e5,
		  loglik = ed1$progress$loglik[21:100020])
plot(ted$iter, ted$loglik, type = "l", xlim = c(0, 1e5), lwd = 2, col = "#D62728", ylab = "log-likelihood", xlab = "iteration")
lines(ed$iter,ed$loglik,type = "l",col = "#17BECF",lwd = 2)

ted_iter = nrow(ted1$progress)
plot_convergence(ted1$progress$loglik[21:ted_iter], ed1$progress$loglik[21:100020], names = c("TED", "ED"), 
                cols = c(c("#D62728", "#17BECF")), xlab = "iteration", ylab = "log10_distance to best log-likelihood", pch = 20, e = 1e-4, cex = 0.3, type = "l", lwd = 2)




#example2
ted <- data.frame(iter = 1: (nrow(ted2$progress)-20),
		  loglik = ted2$progress$loglik[21:nrow(ted2$progress)])
ed  <- data.frame(iter = 1:1e5,
		  loglik = ed2$progress$loglik[21:100020])
plot(ted$iter, ted$loglik, type = "l", xlim = c(0, 1e5), lwd = 2, col = "#D62728", ylab = "log-likelihood", xlab = "iteration")
lines(ed$iter,ed$loglik,type = "l",col = "#17BECF",lwd = 2)

ted_iter = nrow(ted2$progress)
plot_convergence(ted2$progress$loglik[21:ted_iter], ed2$progress$loglik[21:100020], names = c("TED", "ED"), 
                cols = c(c("#D62728", "#17BECF")), xlab = "iteration", ylab = "log10_distance to best log-likelihood", pch = 20, e = 1e-4, cex = 0.3, type = "l", lwd = 2)


#example3
ted <- data.frame(iter = 1: (nrow(ted3$progress)-20),
		  loglik = ted3$progress$loglik[21:nrow(ted3$progress)])
ed  <- data.frame(iter = 1:1e5,
		  loglik = ed3$progress$loglik[21:100020])
plot(ted$iter, ted$loglik, type = "l", xlim = c(0, 1e5), lwd = 2, col = "#D62728", ylab = "log-likelihood", xlab = "iteration")
lines(ed$iter,ed$loglik,type = "l",col = "#17BECF",lwd = 2)


ted_iter = nrow(ted3$progress)
plot_convergence(ted3$progress$loglik[21:ted_iter], ed3$progress$loglik[21:100020], names = c("TED", "ED"), 
                cols = c(c("#D62728", "#17BECF")), xlab = "iteration", ylab = "log10_distance to best log-likelihood", pch = 20, e = 1e-4, cex = 0.3, type = "l", lwd = 2)




#example4
ted <- data.frame(iter = 1: (nrow(ted4$progress)-20),
		  loglik = ted4$progress$loglik[21:nrow(ted4$progress)])
ed  <- data.frame(iter = 1:1e5,
		  loglik = ed4$progress$loglik[21:100020])
plot(ted$iter, ted$loglik, type = "l", xlim = c(0, 1e5), lwd = 2, col = "#D62728", ylab = "log-likelihood", xlab = "iteration")
lines(ed$iter,ed$loglik,type = "l",col = "#17BECF",lwd = 2)


ted_iter = nrow(ted4$progress)
plot_convergence(ted4$progress$loglik[21:ted_iter], ed4$progress$loglik[21:100020], names = c("TED", "ED"), 
                cols = c(c("#D62728", "#17BECF")), xlab = "iteration", ylab = "log10_distance to best log-likelihood", pch = 20, e = 1e-4, cex = 0.3, type = "l", lwd = 2)
#mtext("3(b)", side = 1, line = 5)
```







