---
title: Updated analysis of the GTEx data
author: Yunqi Yang and Peter Carbonetto
output:
  workflowr::wflow_html:
    toc: no
    theme: readable
    highlight: textmate
    lib_dir: site_libs
    self_contained: no
---

Here we summarize the results of the updated udr analysis of the GTEx
data.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the packages used in the analyses below:

```{r load-pkgs, message=FALSE}
library(mashr)
library(ggplot2)
library(cowplot)
```

Load the z-scores:

```{r load-data}
dat <- readRDS("data/data_analysis/GTEx_V8_strong_z.rds")
dim(dat$strong.z)
```

Load the results of the cross-validation:

```{r load-cv}
load("output/data_analysis/gtex_cv_out.RData")
```

These two plots summarize the results of the cross-validation, showing
the setting of $\lambda$ vs. the test-set log-likelihood (relative to
the largest achieved test-set log-likelihood):

```{r cv-results-1, fig.height=2, fig.width=4.5}
n <- 3127
pdat <- data.frame(lambda = log10(LAMBDA),
                   iw = cv_iw_loglik_test/n,
                   nn = cv_nn_loglik_test/n)
pdat <- transform(pdat,
                  iw = iw - max(iw),
				  nn = nn - max(nn))
p1 <- ggplot(pdat,aes(x = lambda,y = iw)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(-1,3,0.5)) +
  labs(x = "log10 lambda",y = "mean test-set loglik") +
  theme_cowplot(font_size = 10)
p2 <- ggplot(pdat,aes(x = lambda,y = nn)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(-1,3,0.5)) +
  labs(x = "log10 lambda",y = "mean test-set loglik") +
  theme_cowplot(font_size = 10)
plot_grid(p1,p2,nrow = 1,ncol = 2)
```

```{r cv-results-ggsave, echo=FALSE}
ggsave("gtex_cv.pdf",
       plot_grid(p1,p2,nrow = 1,ncol = 2),
       height = 2.5,width = 5.2)
```

These are the settings of the penalty strength parameter $\lambda$
that achieve the best test-set log-likelihood for the inverse Wishart
(IW) penalty and the nuclear norm (NN) penalty:

```{r cv-results-2}
LAMBDA[which.max(cv_iw_loglik_test)]
LAMBDA[which.max(cv_nn_loglik_test)]
```

Now load the results from running the different variants of udr on the
training sets and the "left out" test sets:

```{r load-results}
load("output/data_analysis/gtex_main_out.RData")
```

Summarize the results in a table:

```{r summarize-results}
methods <- rownames(loglik_test)
numiter <- matrix(0,10,5)
rownames(numiter) <- methods
colnames(numiter) <- paste0("k",1:5)
for (i in methods) {
  numiter[i,] <- sapply(fits,function (x) nrow(x[[i]]$progress))
}
timings <- timings/60
loglik_test_relative <- rowSums(loglik_test)
loglik_test_relative <- loglik_test_relative - max(loglik_test_relative)
res <- data.frame(ll_test   = loglik_test_relative/n,
                  time_mean = rowMeans(timings),
				  time_min  = apply(timings,1,min),
				  time_max  = apply(timings,1,max),
				  iter_mean = rowMeans(numiter),
				  iter_min  = apply(numiter,1,min),
				  iter_max  = apply(numiter,1,max))
rows <- order(res$ll_test,decreasing = TRUE)
res  <- res[rows,]
round(res,digits = 3)
```

In this table, "ll_test" is the average per-sample test
log-likelihood, and the "iter" columns give the number of updates
performed. The running times are in minutes.

Now load the results from running the EM algorithm on all the data:

```{r load-results-all-data}
load("output/data_analysis/gtex_all_data_out.RData")
```

Compare the running times and number of iterations performed:

```{r examine-results-all-data-1}
timings
nrow(fit_ed$progress)
nrow(fit_ted_iw$progress)
```

Compare the mixture weights:

```{r examine-results-all-data-2, fig.height=2, fig.width=4}
pdat <- data.frame(ed = fit_ed$w,ted_iw = fit_ted_iw$w)
p1 <- ggplot(pdat,aes(x = ed)) +
  geom_histogram(bins = 24,fill = "black",color = "white") +
  scale_y_continuous(breaks = seq(0,20)) +
  theme_cowplot(font_size = 10)
p2 <- ggplot(pdat,aes(x = ted_iw)) +
  geom_histogram(bins = 24,fill = "black",color = "white") +
  scale_y_continuous(breaks = seq(0,20)) +
  theme_cowplot(font_size = 10)
plot_grid(p1,p2,nrow = 1,ncol = 2)
```

```{r results-all-data-ggsave-1, echo=FALSE}
ggsave("gtex_weights.pdf",
       plot_grid(p1,p2,nrow = 1,ncol = 2),
       height = 2,width = 4.2)
```

Now let's take a look at the top effect-sharing patterns from the two
analyses.

```{r effect-plot}
colors <- c("#3288bd","#66c2a5","#abdda4","#e6f598","#ffffbf",
            "#fee08b","#fdae61","#f46d43","#d53e4f")
effect_plot <- function (U, w, k) {
  s <- max(diag(U))
  U <- U/s
  n <- nrow(U)
  plot_title <- sprintf("k%d (w = %0.4f, s = %0.3f)",k,w,s)
  pdat <- cbind(expand.grid(row = 1:n,col = 1:n),
                data.frame(value = as.vector(U)))
  pdat <- subset(pdat,row >= col)
  pdat <- rbind(pdat,data.frame(row = c(1,1),col = c(n-1,n),value = c(-1,1)))
  return(ggplot(pdat,aes(x = col,y = row,fill = value)) +
    geom_tile(color = "white") +
    scale_x_continuous(breaks = c(1,seq(5,n,5),49)) +
    scale_y_continuous(breaks = c(1,seq(5,n,5),49),transform = "reverse") +
    scale_fill_gradientn(colors = colors,values = seq(0,1,length.out = 9)) +
    guides(fill = "none") +
    labs(x = "",y = "",title = plot_title) +
    theme_cowplot(font_size = 8) +
    theme(axis.line = element_blank(),
          plot.title = element_text(size = 8,face = "plain")))
}
```

These are the top 12 effect-sharing patterns from running the ED
updates with no penalty on the covariance matrices:

```{r ed-top-effect-sharing-patterns, fig.height=6.75, fig.width=8}
w <- fit_ed$w
k <- order(fit_ed$w,decreasing = TRUE)
k <- k[1:12]
plots <- vector("list",12)
names(plots) <- paste0("k",k)
for (i in k) {
  U <- fit_ed$U[[i]]$mat
  plots[[paste0("k",i)]] <- effect_plot(U,w[i],i)
}
plot_grid(plots[[1]],plots[[2]],plots[[3]],plots[[4]],
          plots[[5]],plots[[6]],plots[[7]],plots[[8]],
		  plots[[9]],plots[[10]],plots[[11]],plots[[12]],
		  nrow = 3,ncol = 4)
```

```{r ed-top-effect-sharing-patterns-ggsave, echo=FALSE}
ggsave("gtex_sharing_patterns_ed.pdf",
       plot_grid(plots[[1]],plots[[2]],plots[[3]],plots[[4]],
                 plots[[5]],plots[[6]],plots[[7]],plots[[8]],
		         plots[[9]],plots[[10]],plots[[11]],plots[[12]],
		         nrow = 3,ncol = 4),
       height = 7,width = 8)
```

These are the top 12 effect-sharing patterns from running the TED
updates with the IW penalty on the covariance matrices:

```{r ted-iw-top-effect-sharing-patterns, fig.height=6.75, fig.width=8}
w <- fit_ted_iw$w
k <- order(fit_ted_iw$w,decreasing = TRUE)
k <- k[1:12]
plots <- vector("list",12)
names(plots) <- paste0("k",k)
for (i in k) {
  U <- fit_ted_iw$U[[i]]$mat
  plots[[paste0("k",i)]] <- effect_plot(U,w[i],i)
}
plot_grid(plots[[1]],plots[[2]],plots[[3]],plots[[4]],
          plots[[5]],plots[[6]],plots[[7]],plots[[8]],
		  plots[[9]],plots[[10]],plots[[11]],plots[[12]],
		  nrow = 3,ncol = 4)
```

```{r ted-iw-top-effect-sharing-patterns-ggsave, echo=FALSE}
ggsave("gtex_sharing_patterns_ted_iw.pdf",
       plot_grid(plots[[1]],plots[[2]],plots[[3]],plots[[4]],
                 plots[[5]],plots[[6]],plots[[7]],plots[[8]], 
 		         plots[[9]],plots[[10]],plots[[11]],plots[[12]],
		         nrow = 3,ncol = 4),
       height = 7,width = 8)
```

These are the 49 tissues:

```{r tissues}
colnames(dat$strong.z)
```

This is the legend for the effect-sharing plots:

```{r effect-sharing-plot-legend, fig.height=2, fig.width=2.75}
pdat <- data.frame(x = seq(1,3,length.out = 1000),
                   y = seq(1,3,length.out = 1000),
				   value = seq(-1,1,length.out = 1000))
p <- ggplot(pdat,aes(x = x,y = y,color = value)) +
  geom_point() +
  scale_color_gradientn(colors = colors,values = seq(0,1,length.out = 9)) +
  theme_cowplot(font_size = 12)
print(p)
```

```{r effect-sharing-plot-legend-ggsave, echo=FALSE}
ggsave("gtex_sharing_pattern_legend.pdf",p,height = 2,width = 2.75)
```

Let's briefly understand how these choices of prior covariances would
affect a MASH analysis of the GTEx data.

```{r mash, cache=TRUE}
mash_data   <- mashr::mash_set_data(dat$strong.z,V = dat$null.cor)
U_ed        <- lapply(fit_ed$U,"[[","mat")
U_ted_iw    <- lapply(fit_ted_iw$U,"[[","mat")
mash_ed     <- mash(mash_data,U_ed)
mash_ted_iw <- mash(mash_data,U_ted_iw)
```

As expected, using the covariance matrices estimated using the "new"
workflow (smart initialization, TED updates, IW penalty) resulted in
a more even spread of the mixture weights:

```{r mash-weights}
sum(mash_ed$fitted_g$pi > 0.01)
sum(mash_ted_iw$fitted_g$pi > 0.01)
```

This plot compares the 5% quantiles of the lfsrs from the two MASH analyses:

```{r mash-lfsr, fig.height=3, fig.width=3}
res_ed     <- get_lfsr(mash_ed)
res_ted_iw <- get_lfsr(mash_ted_iw)
pdat <- data.frame(ed     = quantile(res_ed,seq(0,1,0.05)),
                   ted_iw = quantile(res_ted_iw,seq(0,1,0.05)))
p <- ggplot(pdat,aes(x = ed,y = ted_iw)) +
  geom_point() +
  geom_line() +
  geom_abline(intercept = 0,slope = 1,color = "magenta",linetype = "dotted") +
  scale_x_continuous(breaks = seq(0,1,0.1),limits = c(0,0.55)) +
  scale_y_continuous(breaks = seq(0,1,0.1),limits = c(0,0.55)) +
  theme_cowplot(font_size = 10)
print(p)
```

The short story here is that the "new" workflow concentrates more of the
lfsrs on values near 1, meaning fewer "significant" effects on gene
expression. I interpret this result as the new workflow providing a
posterior distribution than is better calibrated. (Can we say that the
posterior distribution is better calibrated, or is the lfsrs that are
better calibrated?) The question of how to obtain a "well calibrated"
MASH model remains an open question; one limitation of MASH
(and more generally empirical Bayes methods) is that they do not
fully account for uncertainty in the model and model parameters,
notably in the prior covariance matrices. This is why better estimates
of the prior covariance matrices will result in a more trustworthy
MASH analysis.

```{r mash-lfsr-ggsave, echo=FALSE}
ggsave("gtex_lfsr.pdf",p,height = 2.5,width = 2.5)
```
