---
title: Updated analysis of the GTEx data
author: Yunqi Yang and Peter Carbonetto
output:
  workflowr::wflow_html:
    toc: no
    theme: readable
    highlight: textmate
    lib_dir: site_libs
    self_contained: no
---

Here we summarize the results of the updated udr analysis of the GTEx
data.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load the packages used in the analyses below:

```{r load-pkgs, message=FALSE}
library(ggplot2)
library(cowplot)
```

Load the z-scores:

```{r load-data}
dat <- readRDS("data/data_analysis/GTEx_V8_strong_z.rds")
dim(dat$strong.z)
```

Load the results of the cross-validation:

```{r load-cv}
load("output/data_analysis/gtex_cv_out.RData")
```

These two plots summarize the results of the cross-validation, showing
the setting of $\lambda$ vs. the test-set log-likelihood (relative to
the largest achieved test-set log-likelihood):

```{r cv-results-1, fig.height=2, fig.width=4.5}
n <- 3127
pdat <- data.frame(lambda = log10(LAMBDA),
                   iw = cv_iw_loglik_test/n,
                   nn = cv_nn_loglik_test/n)
pdat <- transform(pdat,
                  iw = iw - max(iw),
				  nn = nn - max(nn))
p1 <- ggplot(pdat,aes(x = lambda,y = iw)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(-1,3,0.5)) +
  labs(x = "log10 lambda",y = "mean test-set loglik") +
  theme_cowplot(font_size = 10)
p2 <- ggplot(pdat,aes(x = lambda,y = nn)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(-1,3,0.5)) +
  labs(x = "log10 lambda",y = "mean test-set loglik") +
  theme_cowplot(font_size = 10)
plot_grid(p1,p2,nrow = 1,ncol = 2)
```

These are the settings of the penalty strength parameter $\lambda$
that achieve the best test-set log-likelihood for the inverse Wishart
(IW) penalty and the nuclear norm (NN) penalty:

```{r cv-results-2}
LAMBDA[which.max(cv_iw_loglik_test)]
LAMBDA[which.max(cv_nn_loglik_test)]
```

Now load the results from running the different variants of udr on the
training sets and the "left out" test sets:

```{r load-results}
load("output/data_analysis/gtex_main_out.RData")
```

Summarize the results in a table:

```{r summarize-results}
methods <- rownames(loglik_test)
numiter <- matrix(0,10,5)
rownames(numiter) <- methods
colnames(numiter) <- paste0("k",1:5)
for (i in methods) {
  numiter[i,] <- sapply(fits,function (x) nrow(x[[i]]$progress))
}
timings <- timings/60
loglik_test_relative <- rowSums(loglik_test)
loglik_test_relative <- loglik_test_relative - max(loglik_test_relative)
res <- data.frame(ll_test   = loglik_test_relative/n,
                  time_mean = rowMeans(timings),
				  time_min  = apply(timings,1,min),
				  time_max  = apply(timings,1,max),
				  iter_mean = rowMeans(numiter),
				  iter_min  = apply(numiter,1,min),
				  iter_max  = apply(numiter,1,max))
rows <- order(res$ll_test,decreasing = TRUE)
res  <- res[rows,]
round(res,digits = 3)
```

TO DO: Re-run (1) TED updates with "smart" initialization and IW
penalty, (2) ED updates with "smart" initialization and no penalty.
Create a new script, gtex_all_samples.R.
