---
title: "Check scale invariant"
output: html_document
date: '2024-04-25'
---

## Description: 

Check scale invariant property from the fitting results. Check for a single scaler $s=5$. 

1. In large n/R + running TED, the scale invariance property satisfies.

2. In smaller n/R + running TED.iw (very strong penalty), there is one matrix that doesn't satisfy scale invariance property. 


```{r}
library(udr)
library(mashr)
library(mvtnorm)
source("/Users/nicholeyang/Downloads/udr-paper/code/simulation_func.R")
```

### 1. Simulate data 

```{r}
set.seed(1)
R = 5
U = list()
U[[1]] = diag(seq(1:5))
U[[2]] = matrix(0, ncol = R, nrow = R)
U[[2]][1,1] = 1
u = rnorm(R)
U[[3]] = u %*% t(u)

# Non-diagonal V
V <- matrix(1, ncol =5, nrow = 5)
diag(V) <- 2

K = length(U)
w = rep(1/K, K)
n_train = 1e4
dat_train = simulate_mixture_ebnm(n_train, w, U, V)
print(U)
```


### Check scale invariant for TED. Large n & small R
```{r}
# fit 1: original fit
maxiter = 2e3
U.init = list()
# The scaler 
s = 5 
for (k in 1:K){
  U.init[[k]] <- udr:::sim_unconstrained(R)
}

f0 = ud_init(dat_train$X, V = V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit1 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                tol = 1e-02, tol.lik = 1e-2, lambda = 0, maxiter = maxiter), verbose=FALSE)

# fit2: scaled fit
f0 = ud_init(s* dat_train$X, V = s^2*V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit2 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                tol = 1e-02, tol.lik = 1e-2, lambda = 0, maxiter = maxiter), verbose=FALSE)

U1 <- lapply(fit1$U,function (e) "[["(e,"mat"))
U2 <- lapply(fit2$U,function (e) "[["(e,"mat"))
```

```{r}
U1
U2
```


### Check scale invariant for TED.iw. Smaller n & strong penalty
```{r   }
##  2. Check scale invariant for TED.iw
# fit 1: original fit
maxiter = 2e3
dat_train = simulate_mixture_ebnm(500, w, U, V)
lambda = 100*R
U.init = list()
for (k in 1:K){
  U.init[[k]] <- udr:::sim_unconstrained(R)
}

f0 = ud_init(dat_train$X, V = V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit1 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                 tol = 1e-02, tol.lik = 1e-2, lambda = lambda, maxiter = maxiter, penalty = "iw"), verbose=FALSE)

# fit 2: scaled fit
U.init = list()
for (k in 1:K){
  U.init[[k]] <- udr:::sim_unconstrained(R)
}

s = 5
f0 = ud_init(s* dat_train$X, V = s^2*V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit2 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                 tol = 1e-02, tol.lik = 1e-2, lambda = lambda, maxiter = maxiter, penalty = "iw"), verbose=FALSE)

U1 <- lapply(fit1$U,function (e) "[["(e,"mat"))
U2 <- lapply(fit2$U,function (e) "[["(e,"mat"))
```

```{r}
U1
```

```{r}
lapply(1:3, function(i) U2[[i]]/25)
```

