---
title: "Check scale invariant"
output: html_document
date: '2024-04-25'
---

## Description: 

Check scale invariant property from the fitting results. Check for a single scaler $s=5$ and a matrix $A$ for single component model. 

```{r}
library(udr)
library(mashr)
library(mvtnorm)
source("/Users/nicholeyang/Downloads/udr-paper/code/simulation_func.R")
```

### 1. Simulate data 

```{r}
set.seed(1)
R = 5
U = list()
U[[1]] = diag(seq(1:5))

# Non-diagonal V
V <- matrix(1, ncol =5, nrow = 5)
diag(V) <- 2

K = length(U)
w = rep(1/K, K)
n_train = 1e4
dat_train = simulate_mixture_ebnm(n_train, w, U, V)
print(U)
```


### 2. Check scale invariant for TED, single scalar. Large n & small R
```{r}
# fit 1: original fit
maxiter = 2e3
U.init = list()

# The scalar 
s = 5 
set.seed(1)
for (k in 1:K){
  U.init[[k]] <- udr:::sim_unconstrained(R)
}

f0 = ud_init(dat_train$X, V = V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit1 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                tol = 1e-02, tol.lik = 1e-2, lambda = 0, maxiter = maxiter), verbose=FALSE)

for (k in 1:K){
  U.init[[k]] <- s^2*U.init[[k]]
}
# fit2: scaled fit
f0 = ud_init(s* dat_train$X, V = s^2*V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit2 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                tol = 1e-02, tol.lik = 1e-2, lambda = 0, maxiter = maxiter), verbose=FALSE)

U1 <- lapply(fit1$U,function (e) "[["(e,"mat"))
U2 <- lapply(fit2$U,function (e) "[["(e,"mat"))
```

```{r}
U1
U2[[1]]/25
```


### 3. Check scale invariant for TED.iw, single scalar. Smaller n & strong penalty
```{r   }
##  2. Check scale invariant for TED.iw
# fit 1: original fit
maxiter = 2e3
dat_train = simulate_mixture_ebnm(500, w, U, V)
lambda = 100*R
U.init = list()
for (k in 1:K){
  U.init[[k]] <- udr:::sim_unconstrained(R)
}

f0 = ud_init(dat_train$X, V = V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit1 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                 tol = 1e-02, tol.lik = 1e-2, lambda = lambda, maxiter = maxiter, penalty = "iw"), verbose=FALSE)

# fit 2: scaled fit
for (k in 1:K){
  U.init[[k]] <- s^2*U.init[[k]]
}

s = 5
f0 = ud_init(s* dat_train$X, V = s^2*V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit2 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                 tol = 1e-02, tol.lik = 1e-2, lambda = lambda, maxiter = maxiter, penalty = "iw"), verbose=FALSE)

U1 <- lapply(fit1$U,function (e) "[["(e,"mat"))
U2 <- lapply(fit2$U,function (e) "[["(e,"mat"))
```

```{r}
U1
U2[[1]]/25
```


### 4. Check scale invariant for TED, a matrix A. Large n & small R
```{r}
# fit 1: original fit
maxiter = 2e3
U.init = list()
# The matrix A 
A = rbind(c(1, 0, 0, 0, 0), c(2, 1, 0, 0, 0))
  
for (k in 1:K){
  U.init[[k]] <- udr:::sim_unconstrained(R)
}

f0 = ud_init(dat_train$X, V = V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit1 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                tol = 1e-02, tol.lik = 1e-2, lambda = 0, maxiter = maxiter), verbose=FALSE)

for (k in 1:K){
  U.init[[k]] <- A %*% U.init[[k]] %*% t(A)
}

# fit2: scaled fit
f0 = ud_init(dat_train$X %*% t(A), V = A %*% V %*% t(A), U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit2 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                tol = 1e-02, tol.lik = 1e-2, lambda = 0, maxiter = maxiter), verbose=FALSE)

U1 <- lapply(fit1$U,function (e) "[["(e,"mat"))
U2 <- lapply(fit2$U,function (e) "[["(e,"mat"))
```

```{r}
A %*% U1[[1]] %*% t(A)
U2
```


### 5. Check scale invariant for TED.iw, a matrix A. Smaller n & strong penalty

```{r}
# fit 1: original fit
maxiter = 2e3
U.init = list()
# The matrix A 
A = rbind(c(1, 0, 0, 0, 0), c(2, 1, 0, 0, 0))

maxiter = 2e3
dat_train = simulate_mixture_ebnm(500, w, U, V)
lambda = 100*R
U.init = list()

for (k in 1:K){
  U.init[[k]] <- udr:::sim_unconstrained(R)
}

f0 = ud_init(dat_train$X, V = V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit1 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                tol = 1e-02, tol.lik = 1e-2, lambda = lambda, maxiter = maxiter, penalty = "iw"), verbose=FALSE)


for (k in 1:K){
  U.init[[k]] <- A %*% U.init[[k]] %*% t(A)
}

# fit2: scaled fit
f0 = ud_init(dat_train$X %*% t(A), V = A %*% V %*% t(A), U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
fit2 = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                tol = 1e-02, tol.lik = 1e-2, lambda = lambda, maxiter = maxiter, penalty = "iw"), verbose=FALSE)

U1 <- lapply(fit1$U,function (e) "[["(e,"mat"))
U2 <- lapply(fit2$U,function (e) "[["(e,"mat"))
```

```{r}
A %*% U1[[1]] %*% t(A)
U2
```


