---
title: "test_tpr_code"
author: "Yunqi Yang"
date: "10/17/2022"
output: html_document
---

### Description:

Change sample size and see how true positive rate change. Threshold is chosen to be $0.95$ and we consider the sign of estimated effects. If the estimated effects (non-null) have different sign as true theta, it is not true positive. 

1. Simulation: True $U$s contains 5 singleton matrix + 5 unconstrained matrix. 


```{r knitr, echo=FALSE}
knitr::opts_chunk$set(comment = "#",results = "hold",collapse = TRUE,
                      fig.align = "center", warning = FALSE, message = FALSE)
```

```{r}
source("./code/simulation_func.R")
library(udr)
library(mashr)
library(mvtnorm)
library(LaplacesDemon)
library(ggplot2)
```

```{r }
#### True param in simulation
K = 10
R = 5
s = 5
n.train = c(1e2, 5e2, 1e3, 5e3)
threshold = 0.95
param = list(w = rep(1/K, K), U = list(), V = diag(R))
```

## Simulation: True $U$s contains 5 singleton matrix + 5 unconstrained matrix. 
```{r  cache = TRUE}
#### Simulation and fitting
seeds = c(1:20)

oracle <- matrix(0, ncol = 4, nrow = length(seeds))
ed1 <- matrix(0, ncol = 4, nrow = length(seeds))
ted1 <- matrix(0, ncol = 4, nrow = length(seeds))
ed2 <- matrix(0, ncol = 4, nrow = length(seeds))
ted2 <- matrix(0, ncol = 4, nrow = length(seeds))
fa <- matrix(0, ncol = 4, nrow = length(seeds))


for (i in 1:length(n.train)){
  for (seed in seeds){
  # Simulate data
      set.seed(seed)
      param$U <-  sim_U_true(R, s, null.mat = FALSE, identity = TRUE, cov_structured = FALSE,
                             num_singleton = 5, num_unconstrained = 4)
      dat.train <- simulate_mixture_ebnm(n.train[i], param$w,param$U,param$V)
      mc <- mash_set_data(dat.train$X, Shat = 1)
      ## Oracle tpr
      g.fitted <- create_g(param$w, param$U)
      oracle[seed, i] = oracle[seed, i] + compare_lfsr_fitted_vs_truth(dat.train, g.fitted, mc, threshold = threshold)[[1]]
      
      # Initialization
      set.seed(888)
      U.init = c()
      for (k in 1:K){
        U.init[[k]] <- udr:::sim_unconstrained(R)
      }
      f0 = ud_init(X = dat.train$X, V = param$V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)
      
      # Fit ED
      fit = ud_fit(f0, control = list(unconstrained.update = "ed", resid.update = 'none',
                                      tol = 1e-02, tol.lik = 1e-2, n0 = 0, lambda = 0, maxiter = 5e3), verbose=FALSE)
      U <- lapply(fit$U,function (e) "[["(e,"mat"))
      g.fitted <- create_g(fit$w, U)
      ed1[seed, i] = ed1[seed, i] + compare_lfsr_fitted_vs_truth(dat.train, g.fitted, mc, threshold = threshold)[[1]]
  
      # Fit TED
      fit = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                      tol = 1e-02, tol.lik = 1e-2, n0 = 0, lambda = 0, maxiter = 5e3), verbose=FALSE)
      
      U <- lapply(fit$U,function (e) "[["(e,"mat"))
      g.fitted <- create_g(fit$w, U)
      ted1[seed, i] = ted1[seed, i] + compare_lfsr_fitted_vs_truth(dat.train, g.fitted, mc, threshold = threshold)[[1]]
  
      # Fit ED.reg
      fit = ud_fit(f0, control = list(unconstrained.update = "ed", resid.update = 'none',
                                      tol = 1e-02, tol.lik = 1e-2, n0 = round(R/2), lambda = 0, maxiter = 5e3), verbose=FALSE)
      
      U <- lapply(fit$U,function (e) "[["(e,"mat"))
      g.fitted <- create_g(fit$w, U)
      ed2[seed, i] = ed2[seed, i] + compare_lfsr_fitted_vs_truth(dat.train, g.fitted, mc, threshold = threshold)[[1]]

  
      # Fit TED.reg
      fit = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                      tol = 1e-02, tol.lik = 1e-2, n0 = 0, lambda = round(R/2), maxiter = 5e3), verbose=FALSE)
      U <- lapply(fit$U,function (e) "[["(e,"mat"))
      g.fitted <- create_g(fit$w, U)
      ted2[seed, i] = ted2[seed, i] + compare_lfsr_fitted_vs_truth(dat.train, g.fitted, mc, threshold = threshold)[[1]]
    }
}


```

```{r}
par(mfrow = c(2,2))
boxplot(oracle, names = n.train, xlab = "n.train", ylab = "true positive rate", main = "Oracle")
boxplot(ed1, names = n.train, xlab = "n.train", ylab = "true positive rate", main = "ED")
boxplot(ted1, names = n.train, xlab = "n.train", ylab = "true positive rate", main = "TED")
boxplot(ed2, names = n.train, xlab = "n.train", ylab = "true positive rate", main = "ED.reg")
#boxplot(ted2, names = n.train, xlab = "n.train", ylab = "true positive rate", main = "TED.reg")

```

## Plot the distribution of thetas under different sample sizes. One data example.

```{r}
set.seed(2)
dat.train = list()
param$U <-  sim_U_true(R, s, null.mat = FALSE, identity = TRUE, num_singleton = 5, num_unconstrained = 4)

for (i in 1:length(n.train)){
  dat.train[[i]] <- simulate_mixture_ebnm(n.train[i], param$w,param$U,param$V)
}
```

```{r}
par(mfrow = c(2,2))
for (i in 1:length(n.train)){
  hist(dat.train[[i]]$theta, xlab = "theta", main = paste0("n = ", n.train[i]), breaks = 20, freq = FALSE)
}
```

#### The proportion of strong signals (defined as abs(theta) > 1):
```{r}
strong_signals <- rep(NA, length(n.train))
for (i in 1:length(n.train)){
  num_effects = n.train[i] * R
  strong_signals[i] = sum(abs(dat.train[[i]]$theta) > 1)/ num_effects
}
strong_signals
```

## Plot for true theta: n = 100
```{r}
df <- expand.grid(x = 0:4, y = 0:99)
df$z <- c(t(dat.train[[1]]$theta))
ggplot(df, aes(x = x, y = y, fill = z)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "#075AFF",
                       mid = "#FFFFCC",
                       high = "#FF0000")
```

## Plot for true theta: n = 500
```{r}

df <- expand.grid(x = 0:4, y = 0:499)
df$z <- c(t(dat.train[[2]]$theta))
ggplot(df, aes(x = x, y = y, fill = z)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "#075AFF",
                       mid = "#FFFFCC",
                       high = "#FF0000")
```


