---
title: "Regularized fitting varying components"
author: "Yunqi Yang"
date: "11/17/2022"
output: html_document
---

#### Description: 

This file try to assess why regularized methods is stable across different number of components. I compare the patterns between fitted Us (top 6 patterns) and the truth.

The truth contains 6 different patterns. First, I initialize 6 components for TED.reg, ED.reg and TED. Then I initialize 100 components for all the algorithms: TED.reg, ED.reg and TED. 

```{r include = FALSE}
knitr::opts_chunk$set(comment = "#",
                      results = "hold",
                      collapse = TRUE,
                      fig.align = "center",
                      warning = FALSE,
                      message = FALSE)
```


```{r}
source("./code/simulation_func.R")
library(udr)
library(mashr)
library(mvtnorm)
library(LaplacesDemon)
library(ggplot2)
require(gridExtra)
library(cowplot)
```

```{r}
#### True param in simulation
K = 6
R = 50
s = 5
n.train = 1e3
param = list(w = rep(1/K, K), U = list(), V = diag(R))
```

```{r}
set.seed(4)
# True Us: 1 singleton + equal_effect + identity + 3 partial shared + 4 rinvWishart.
param$U <- sim_U_true(R, s, identity = TRUE, partial_sharing = TRUE,
                       equal_effect = TRUE, num_singleton = 1)
X.train <- simulate_ud_data(n.train, param$w,param$U,param$V)
```

## True patterns
```{r}
plot.true <- list()
for (i in 1:6) {
  df <- expand.grid(x = 0:49, y = 0:49)
  df$z <- c(t(param$U[[i]]))
  p <- ggplot(df, aes(x = x, y = y, fill = z)) + geom_tile(color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                        mid = "#FFFFCC",
                        high = "#FF0000")
  plot.true[[i]] <- p 
}
```

```{r fig.width = 8, fig.height = 12}
plot_grid(plot.true[[1]], plot.true[[2]], plot.true[[3]], 
          plot.true[[4]], plot.true[[5]], plot.true[[6]],
          ncol = 2, byrow = TRUE)
```

# Fit 6 components 
```{r cache = TRUE}

# Initialization
ncomp = 6
set.seed(888)
U.init = c()
for (k in 1:ncomp){
  U.init[[k]] <- udr:::sim_unconstrained(R)
}

f0 = ud_init(X = X.train, V = param$V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)

fit.ted_reg = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                        tol = 1e-02, tol.lik = 1e-2, n0 = 0, lambda = round(R), maxiter = 5e3), verbose=FALSE)

fit.ed_reg = ud_fit(f0, control = list(unconstrained.update = "ed", resid.update = 'none',
                                    tol = 1e-02, tol.lik = 1e-2, n0 = round(R), lambda = 0, maxiter = 5e3), verbose=FALSE)
fit.ted = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                        tol = 1e-02, tol.lik = 1e-2, n0 = 0, lambda = 0, maxiter = 5e3), verbose=FALSE)
```

#### Check mixture model weights
```{r }
fit.ted_reg$w 
fit.ed_reg$w 
fit.ted$w
```

```{r}
U.ted_reg <- lapply(fit.ted_reg$U,function (e) "[["(e,"mat"))
U.ed_reg <- lapply(fit.ed_reg$U,function (e) "[["(e,"mat"))
U.ted <- lapply(fit.ed_reg$U,function (e) "[["(e,"mat"))
```

```{r}
ted_reg <- list() 
ed_reg <- list()
ted <- list()

for (i in 1:6) {
  U <- U.ted_reg[[i]]
  df <- expand.grid(x = 0:49, y = 0:49)
  df$z <- c(t(U))
  p <- ggplot(df, aes(x = x, y = y, fill = z)) + geom_tile(color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                        mid = "#FFFFCC",
                        high = "#FF0000")
  ted_reg[[i]] <- p 
}

for (i in 1:6) {
  U <- U.ed_reg[[i]]
  df <- expand.grid(x = 0:49, y = 0:49)
  df$z <- c(t(U))
  p <- ggplot(df, aes(x = x, y = y, fill = z)) + geom_tile(color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                        mid = "#FFFFCC",
                        high = "#FF0000")
  ed_reg[[i]] <- p 
}

for (i in 1:6) {
  U <- U.ted[[i]]
  df <- expand.grid(x = 0:49, y = 0:49)
  df$z <- c(t(U))
  p <- ggplot(df, aes(x = x, y = y, fill = z)) + geom_tile(color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                        mid = "#FFFFCC",
                        high = "#FF0000")
  ted[[i]] <- p 
}

```

## TED.reg patterns from 6 components fit
```{r fig.width = 8, fig.height = 12}
plot_grid(ted_reg[[1]], ted_reg[[2]], ted_reg[[3]],
          ted_reg[[4]], ted_reg[[5]], ted_reg[[6]],
          ncol = 2, byrow = TRUE)
```


## ED.reg patterns from 6 components fit
```{r fig.width = 8, fig.height = 12}
plot_grid(ed_reg[[1]], ed_reg[[2]], ed_reg[[3]],
          ed_reg[[4]], ed_reg[[5]], ed_reg[[6]],
          ncol = 2, byrow = TRUE)
```


## TED patterns from 6 components fit
```{r fig.width = 8, fig.height = 12}
plot_grid(ted[[1]], ted[[2]], ted[[3]],
          ted[[4]], ted[[5]], ted[[6]],
          ncol = 2, byrow = TRUE)
```


# 100 components
```{r cache = TRUE}
# Initialization
set.seed(888)
U.init = c()
for (k in 1:ncomp){
  U.init[[k]] <- udr:::sim_unconstrained(R)
}

f0 = ud_init(X = X.train, V = param$V, U_scaled = NULL, U_unconstrained = U.init, n_rank1 = 0)

fit.ted_reg = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                        tol = 1e-02, tol.lik = 1e-2, n0 = 0, lambda = round(R), maxiter = 5e3), verbose=FALSE)
fit.ed_reg = ud_fit(f0, control = list(unconstrained.update = "ed", resid.update = 'none',
                                    tol = 1e-02, tol.lik = 1e-2, n0 = round(R), lambda = 0, maxiter = 5e3), verbose=FALSE)
fit.ted = ud_fit(f0, control = list(unconstrained.update = "ted", resid.update = 'none',
                                        tol = 1e-02, tol.lik = 1e-2, n0 = 0, lambda = 0, maxiter = 5e3), verbose=FALSE)
```

#### Check mixture model weights
```{r}
sum(fit.ted_reg$w > 1e-2)
sum(fit.ed_reg$w > 1e-2)
sum(fit.ted$w > 1e-2)
```

```{r}
U.ted_reg <- lapply(fit.ted_reg$U,function (e) "[["(e,"mat"))
U.ed_reg <- lapply(fit.ed_reg$U,function (e) "[["(e,"mat"))
U.ted <- lapply(fit.ed_reg$U,function (e) "[["(e,"mat"))

indx1 = order(fit.ted_reg$w, decreasing = TRUE)[1:6]
indx2 = order(fit.ed_reg$w, decreasing = TRUE)[1:6]
indx3 = order(fit.ted$w, decreasing = TRUE)[1:6]
```

```{r}
ted_reg <- list() 
ed_reg <- list()
ted <- list()

for (i in 1:6) {
  U <- U.ted_reg[[indx1[i]]]
  df <- expand.grid(x = 0:49, y = 0:49)
  df$z <- c(t(U))
  p <- ggplot(df, aes(x = x, y = y, fill = z)) + geom_tile(color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                        mid = "#FFFFCC",
                        high = "#FF0000")
  ted_reg[[i]] <- p 
}

for (i in 1:6) {
  U <- U.ed_reg[[indx2[i]]]
  df <- expand.grid(x = 0:49, y = 0:49)
  df$z <- c(t(U))
  p <- ggplot(df, aes(x = x, y = y, fill = z)) + geom_tile(color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                        mid = "#FFFFCC",
                        high = "#FF0000")
  ed_reg[[i]] <- p 
}

for (i in 1:6) {
  U <- U.ted[[indx3[i]]]
  df <- expand.grid(x = 0:49, y = 0:49)
  df$z <- c(t(U))
  p <- ggplot(df, aes(x = x, y = y, fill = z)) + geom_tile(color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                        mid = "#FFFFCC",
                        high = "#FF0000")
  ted[[i]] <- p 
}

```

## TED.reg patterns top 6 from 100 components
```{r fig.width = 8, fig.height = 12}
plot_grid(ted_reg[[1]], ted_reg[[2]], ted_reg[[3]],
          ted_reg[[4]], ted_reg[[5]], ted_reg[[6]],
          ncol = 2, byrow = TRUE)
```


## ED.reg patterns top 6 from 100 components
```{r fig.width = 8, fig.height = 12}
plot_grid(ed_reg[[1]], ed_reg[[2]], ed_reg[[3]],
          ed_reg[[4]], ed_reg[[5]], ed_reg[[6]],
          ncol = 2, byrow = TRUE)
```


## TED patterns top 6 from 100 components
```{r fig.width = 8, fig.height = 12}
plot_grid(ted[[1]], ted[[2]], ted[[3]],
          ted[[4]], ted[[5]], ted[[6]],
          ncol = 2, byrow = TRUE)
```


